---
title: "Chap_02"
author: "Leticia"
date: '2022-09-15'
output: html_document
---
#Libraries
```{r setup, include=FALSE}
library(fpp3)
library(lubridate) #use period() for more complicated datasets - more than 1 ts column
library(GGally)
```

#2.1 Data - tsibble

Create a tsibbel object for time series, remember to set up the index, the time series variable.
```{r}
y <- tsibble(Year = 2015 : 2019,
             Observation = c(123, 39, 78, 52, 110),
             index = Year)

```


##Olympic Running
```{r}
olympic_running %>% distinct(Sex) #Shows the categories
```

##Monthly dataset
```{r}
a10 <- PBS %>% 
  filter(ATC2 == "A10") %>% #Rows
  select(Month, Concession, Type, Cost) %>% #Columns 
  summarise(TotalC = sum(Cost)) %>% 
  mutate(Cost = TotalC/1e6) 

```

##Reading csv turn into tsibble
```{r}
prison <- readr::read_csv("https://OTexts.com/fpp3/extrafiles/prison_population.csv")
prison <- prison %>%
  mutate(Quarter = yearquarter(Date)) %>%
  select(-Date) %>%
  as_tsibble(key = c(State, Gender, Legal, Indigenous),
             index = Quarter)

prison
```


#2.2 Plot - autoplot()
##melsyd_economy plot
```{r}
melsyd_economy <- ansett %>%
   filter(Airports == "MEL-SYD", Class == "Economy") %>%
   mutate(Passengers = Passengers/1000)
 autoplot(melsyd_economy, Passengers) + #autoplot is created for time series
   labs(title = "Ansett airlines economy class",
        subtitle = "Melbourne-Sydney",
        y = "Passengers ('000)")


melsyd_economy
```
##a10 plot
```{r}
autoplot(a10, Cost)+
  labs(y= "$ (millions)",
       title = "Australian antidiabetic drugs sales")
```

#2.3 Time series patterns

Patterns can be: Trend, seasonal and Cyclic.

## ggseason
```{r}
a10 %>%
  gg_season(Cost, labels = "both") + #note 
  labs(y = "$ (millions)",
       title = "Seasonal plot: Antidiabetic drug sales")
```

##vic_elec
```{r}
vic_elec %>% gg_season(Demand, period = "day") +
   theme(legend.position = "none") #+
#   labs(y="MWh", title="Electricity demand: Victoria")

```
```{r}
vic_elec %>% gg_season(Demand, period = "week") +
  theme(legend.position = "none") +
  labs(y="MWh", title="Electricity demand: Victoria")
```

```{r}
vic_elec %>% gg_season(Demand, period = "year") +
  labs(y="MWh", title="Electricity demand: Victoria")
```
#2.5 Subseries
```{r}
a10 %>%
  gg_subseries(Cost) + #subseries already identified the year and month
  labs(
    y = "$ (millions)",
    title = "Australian antidiabetic drug sales"
  )

#blue lines - means
```
##tourism
```{r}
holidays <- tourism %>%
  filter(Purpose == "Holiday") %>% 
  group_by(State) %>%
  summarise(Trips = sum(Trips))
  
  holidays
```
##autoplot tourism
```{r}
autoplot(holidays, Trips) +
  labs(y = "Overnight trips ('000)",
       title = "Australian domestic holidays")
```
##season tourism
```{r}
gg_season(holidays, Trips) +
  labs(y = "Overnight trips ('000)",
       title = "Australian domestic holidays")
```
##Series for tourism
```{r}
holidays %>%
  gg_subseries(Trips) +
  labs(y = "Overnight trips ('000)",
       title = "Australian domestic holidays")
```
#2.6 Scatterplots
```{r}
vic_elec %>%
  filter(year(Time) == 2014) %>%
  autoplot(Demand) +
  labs(y = "GW",
       title = "Half-hourly electricity demand: Victoria")

```

```{r}
vic_elec %>%
  filter(year(Time) == 2014) %>%
  autoplot(Temperature) +
  labs(
    y = "Degrees Celsius",
    title = "Half-hourly temperatures: Melbourne, Australia"
  )
```
##Correlation Plot
```{r}
vic_elec %>%
  filter(year(Time) == 2014) %>%
  ggplot(aes(x = Temperature, y = Demand)) +
  geom_point() +
  labs(x = "Temperature (degrees Celsius)",
       y = "Electricity demand (GW)")
```
##Scatterplot Matrices
```{r}
visitors <- tourism %>%
  group_by(State) %>%
  summarise(Trips = sum(Trips))
visitors %>%
  ggplot(aes(x = Quarter, y = Trips)) +
  geom_line() +
  facet_grid(vars(State), scales = "free_y") +
  labs(title = "Australian domestic tourism",
       y= "Overnight trips ('000)")
```

##Relationships
```{r}
visitors %>%
  pivot_wider(values_from=Trips, names_from=State) %>%
  GGally::ggpairs(columns = 2:9)

```

#2.7 Lag plots
```{r}
recent_production <- aus_production %>%
  filter(year(Quarter) >= 2000)
recent_production %>%
  gg_lag(Beer, geom = "point", lags = 1:16) +
  labs(x = "lag(Beer, k)")
```

#2.8 Autocorrelation

If X, Y are two sets of numbers with a 1-1 correspondence between them, then we define the $$cor(X,Y)=\frac{\sum (x-\overline{x})(y-\overline{y})}{\sqrt{\sum (x-\overline{x})^2 * \sum(y-\overline{y})^2}}$$


ACF is a function to calculate the correlation between lagged values of a time series. 

```{r}
recent_production %>% ACF(Beer, lag_max = 9) #acf autocorrelation function
```
```{r}
recent_production %>%
  ACF(Beer) %>%
  autoplot() + labs(title="Australian beer production")
```
#2.9 White noise

This looks like a residual plot. Those are the NO AUTOCORRELATION.

```{r}
set.seed(30)
y <- tsibble(sample = 1:50, wn = rnorm(50), index = sample) #create a new data set with random numbers, index is also being determined.
y %>% autoplot(wn) + labs(title = "White noise", y = "")
```

```{r}
y %>%
  ACF(wn) %>%
  autoplot() + labs(title = "White noise")
```
The black lines doesn't touch the blue dashed lines and the ACF lies within ±2/√T where T is the length of the time. T=50 in this case = ±0.28. If its more than 5%, it could not be white noise.
2.10 Exercise 13

```{r}
l <- gafa_stock
levels(as.factor(l$Symbol))

```


```{r}
dgoog <- gafa_stock %>%
  filter(Symbol == "GOOG", year(Date) >= 2018) %>%
  mutate(trading_day = row_number()) %>%
  update_tsibble(index = trading_day, regular = TRUE) %>%
  mutate(diff = difference(Close))

dgoog
```
##a) 
The code was changed from date to a sequential number.

##b) ACF
```{r}
dgoog %>% ACF(diff, lag_max = 100) %>% 
  autoplot() +
  labs(title="One day difference closing") #monthly?


```
##c) white noise?

Yes, more than 95% of the black lines is inside the blue dashed lines and ±2/√T= ± 0.57. But I dont know if T is corrected.

```{r}
a=2/sqrt(12)
a
```


```{r}
dgoog %>%
  ACF(diff) %>%
  autoplot() + labs(title="Google Stocks")
```


#2.12 Questions
sibble c
This tontains 64 separate time series corresponding to the combinations of the 8 states, 2 genders, 2 legal statuses and 2 indigenous statuses. Each of these series is 48 observations in length, from 2005 Q1 to 2016 Q4. #I dont understand how they come up with this numbers.

Is the autoplot created based on ggplot2? Yes, 100%. Shortcut for timeseries.

I don't get difference between cyclic and seasonal. 2.3 got it

vic_elec has this type of data s3: posix

How to decide the type of pattern? Based on researcher knowledge? 2.5 a10 example

I dont see where holiday is in the tourism dataset - Got it!

I didn't understand the Lag plot - Lagged in production? why is this important? From this on I dont get it.

Lag_max =9 is the number of Quarters? what is k? I dont understand the correlation graph.

I still can not see the prediction, should we supose to know by now?


















