---
title: "Chapter 3 Time series decomposition"
author: "Leticia"
date: '2022-10-06'
output: html_document
---

Analyse trend and cycles 

  - Calendar adjustments
  - Population adjustments
  - Inflation adjustments
  - Mathematical transformations

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fpp3)
library(slider)
```

# Transformation of GDP to GDP per/capita
## Population adjustments
```{r cars}
global_economy %>%
  filter(Country == "Australia") %>%
  autoplot(GDP/Population) +
  labs(title= "GDP per capita", y = "$US")
```
##  Inflation adjustments
```{r}
print_retail <- aus_retail %>%
  filter(Industry == "Newspaper and book retailing") %>%
  group_by(Industry) %>%
  index_by(Year = year(Month)) %>%
  summarise(Turnover = sum(Turnover)) #The summation is by year

print_retail
```
```{r}
aus_economy <- global_economy %>%
  filter(Code == "AUS")
aus_economy
```

```{r}
print_retail %>%
  left_join(aus_economy, by = "Year") %>% #joining with the other table by year
  mutate(Adjusted_turnover = Turnover / CPI * 100) %>% #Calculating the turnover (volume de negocios)
  pivot_longer(c(Turnover, Adjusted_turnover),
               values_to = "Turnover") %>% #IDK what this is doing
  mutate(name = factor(name,
         levels=c("Turnover","Adjusted_turnover"))) %>% #has the same value - turnover and adj_turnover, why? Not really, plot will show
  ggplot(aes(x = Year, y = Turnover)) +
  geom_line() +
  facet_grid(name ~ ., scales = "free_y") +
  labs(title = "Turnover: Australian print media industry",
       y = "$AU")
```
```{r}
lambda <- aus_production %>%
  features(Gas, features = guerrero) %>%
  pull(lambda_guerrero) #This ill choose lambda
aus_production %>%
  autoplot(box_cox(Gas, lambda)) +
  labs(y = "",
       title = latex2exp::TeX(paste0(
         "Transformed gas production with $\\lambda$ = ",
         round(lambda,2))))
```
# 3.2 Time Series components

Example: Employment in the US retail sector

```{r}
us_retail_employment <- us_employment %>%
  filter(year(Month) >= 1990, Title == "Retail Trade") %>%
  select(-Series_ID)
autoplot(us_retail_employment, Employed) +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail")
```

```{r}
dcmp <- us_retail_employment %>%
  model(stl = STL(Employed))
components(dcmp) #Dable - a decomposition table

```

```{r}
components(dcmp) %>%
  as_tsibble() %>%
  autoplot(Employed, colour="gray") +
  geom_line(aes(y=trend), colour = "#D55E00") + #the y=trend is to add the orange line? Yes. Nice!
  labs(
    y = "Persons (thousands)",
    title = "Total employment in US retail"
  )
```

## Plot all the components by the time frame
```{r}
components(dcmp) %>% autoplot() #This code understand to plot all the variables in just one plot, it didn't need the facet
```
## Seasonally adjusted data
```{r}
components(dcmp) %>%
  as_tsibble() %>%
  autoplot(Employed, colour = "gray") +
  geom_line(aes(y=season_adjust), colour = "#0072B2") + #another type of adjustment
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail")
```

# 3.3 Moving averages
 
m-MA, Moving average in order of m. It considers the 2 values after and two values before the average that you want.

This is smooth the curves in the graphs. It takes a period of time and returns the mean for this period (season). 

```{r}
global_economy %>%
  filter(Country == "Australia") %>%
  autoplot(Exports) +
  labs(y = "% of GDP", title = "Total Australian exports")

```
## "Slidding" 5-MA☺ 
```{r}
aus_exports <- global_economy %>%
  filter(Country == "Australia") %>%
  mutate(
    `5-MA` = slider::slide_dbl(Exports, mean,
                .before = 2, .after = 2, .complete = TRUE) # It can be 5,7,9 ... assumption: m = 2k+1
  ) #Creating de column with th moving average calculation
```


```{r}
aus_exports %>%
  autoplot(Exports) +
  geom_line(aes(y = `5-MA`), colour = "#D55E00") + #Is this using a column or it's a formula of seasonal behavior?
  labs(y = "% of GDP",
       title = "Total Australian exports") +
  guides(colour = guide_legend(title = "series")) #It excludes the missing values - first to and last two
```

## Moving averages of moving averages
```{r}
beer <- aus_production %>%
  filter(year(Quarter) >= 1992) %>%
  select(Quarter, Beer)
beer_ma <- beer %>%
  mutate(
    `4-MA` = slider::slide_dbl(Beer, mean,
                .before = 1, .after = 2, .complete = TRUE),
    `2x4-MA` = slider::slide_dbl(`4-MA`, mean,
                .before = 1, .after = 0, .complete = TRUE)
  )
```
## Estimating the trend-cycle with seasonal data

Example: Employment in the US retail sector

```{r}
us_retail_employment_ma <- us_retail_employment %>%
  mutate(
    `12-MA` = slider::slide_dbl(Employed, mean,
                .before = 5, .after = 6, .complete = TRUE),
    `2x12-MA` = slider::slide_dbl(`12-MA`, mean,
                .before = 1, .after = 0, .complete = TRUE)
  )
us_retail_employment_ma %>%
  autoplot(Employed, colour = "gray") +
  geom_line(aes(y = `2x12-MA`), colour = "#D55E00") +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail")
```
# 3.4 Classical decomposition

Assumption: The seasonal component is constant from year to year. # No Covid data included.

## Additive decomposition
```{r}
us_retail_employment %>%
  model(
    classical_decomposition(Employed, type = "additive")
  ) %>%
  components() %>%
  autoplot() +
  labs(title = "Classical additive decomposition of total
                  US retail employment")

```
## Multiplicative decomposition
Change the subtraction for divisions.

# 3.5 Methods used by official statistics agencies
STL - classical decomposition (additive)
X_13ARIMA_SEATS multiplicative decomposition

##X_11

```{r}
x11_dcmp <- us_retail_employment %>%
  model(x11 = X_13ARIMA_SEATS(Employed ~ x11())) %>% #X_!#ARIMA_SEATS is the decomposition method - x11()?
  components()
autoplot(x11_dcmp) +
  labs(title =
    "Decomposition of total US retail employment using X-11.")
```
US retail employment: the original data (grey), the trend-cycle component (orange) and the seasonally adjusted data (barely visible in blue).
```{r}
x11_dcmp %>%
  ggplot(aes(x = Month)) +
  geom_line(aes(y = Employed, colour = "Data")) +
  geom_line(aes(y = season_adjust,
                colour = "Seasonally Adjusted")) +
  geom_line(aes(y = trend, colour = "Trend")) +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail") +
  scale_colour_manual(
    values = c("gray", "#0072B2", "#D55E00"),
    breaks = c("Data", "Seasonally Adjusted", "Trend")
  )
```

Seasonal sub-series plot of the seasonal component from the X-11 method applied to total US retail employment.
```{r}
x11_dcmp %>%
  gg_subseries(seasonal)
```

## SEATS method
Seasonal Extraction in ARIMA Time Series

```{r}
seats_dcmp <- us_retail_employment %>%
  model(seats = X_13ARIMA_SEATS(Employed ~ seats())) %>%
  components()
autoplot(seats_dcmp) +
  labs(title =
    "Decomposition of total US retail employment using SEATS")
```

# 3.6 STL decomposition
Seasonal and Trend decomposition using Loess - Here you have more control of the parameters, you can actually set up values for season and cycles.
```{r}
us_retail_employment %>%
  model(
    STL(Employed ~ trend(window = 7) +
                   season(window = "periodic"),
    robust = TRUE)) %>%
  components() %>%
  autoplot()
```


 
# Exersices 
## 2. For each of the following series, make a graph of the data. If transforming seems appropriate, do so and describe the effect.

- United States GDP from global_economy.

```{r Population adjustments}
US_GDP <- global_economy %>% 
  filter(Country == "United States") %>% 
  select(c(Country, Year, Population, GDP)) %>% 
autoplot(GDP/Population) + 
  labs(x="Year", y="GPD per capita")
US_GDP
```
- Slaughter of Victorian “Bulls, bullocks and steers” in aus_livestock
```{r}
slaughter <- aus_livestock %>% 
  filter(Animal=="Bulls, bullocks and steers") %>% 
  group_by(Animal) %>%
  index_by(Year = year(Month)) %>%
  summarise(Count = sum(Count)) %>% 
    autoplot(Count/1000) + labs (y= "Bulls, bullocks and steers (x1000)",
                                 x= "Years")
slaughter
```
- Victorian Electricity Demand from vic_elec
```{r}
vic <- vic_elec %>% 
  index_by(Date) %>%
  summarise(Demand = sum(Demand)) %>% 
  model(stl = STL(Demand)) %>% 
  components() %>% 
  autoplot()
vic
```


- Gas production from aus_production
```{r}
gas_prod <- aus_production %>% 
  index_by(Quarter) %>%
  summarise(Gas = sum(Gas)) %>% 
  model(stl = STL(Gas)) %>% 
  components() %>% 
  autoplot()
gas_prod
  
  
```

## 7. Consider the last five years of the Gas data from aus_production.


7.1 Plot the time series. Can you identify seasonal fluctuations and/or a trend-cycle?
7.2 Use classical_decomposition with type=multiplicative to calculate the trend-cycle and seasonal indices.
7.3 Do the results support the graphical interpretation from part a?
Yes.

```{r}
gas <- tail(aus_production, 5*4) %>% select(Gas) %>% 
  model(classical_decomposition(Gas, type = "multiplicative")) %>%
  components() %>%
  autoplot() +
  labs(title = "Gas prodution in Australia")

gas
```
7.4 Compute and plot the seasonally adjusted data.

Adjust with what parameter?

```{r}
lambda <- aus_production %>%
  features(Gas, features = guerrero) %>%
  pull(lambda_guerrero)

gas <- tail(aus_production, 5*4) %>% select(Gas) %>%
  model(classical_decomposition(Gas, type = "multiplicative")) %>%
  components() %>%
  autoplot(box_cox(Gas, lambda)) + # Adjusting here
  labs(y = "",
       title = latex2exp::TeX(paste0(
         "Transformed gas production with $\\lambda$ = ",
         round(lambda,2))))
  
gas
```


7.6 Change one observation to be an outlier (e.g., add 300 to one observation), and recompute the seasonally adjusted data. What is the effect of the outlier?

```{r}
new_row <- c("300", "2005 Q2")
gas_out <- tail(aus_production, 5*4) %>% select(Gas)
aus_production[1,"Gas"]=300
#Chapter two explains how to add the outlier 
gas_out_new <- rbind(gas_out, new_row)  
```

```{r}
data2 <- gas_out                              # Replicate data
data2[nrow(gas_out) + 1, ] <- new_row         # Add new row
data2 
```

7.7 Does it make any difference if the outlier is near the end rather than in the middle of the time series?

## Questions
1. What is pivot_longer doing (line 47)
2. Formula of BoxCox
3. What is tsibble? A table? (line 90)
4. Does this transformations are based/same on the exponential regression? 
5. Not sure about the remainder graph
6. Is this using a column or it's a formula of seasonal behavior? line(137)
7. I am not sure about the difference between Multiplicative and additive decomposition in a practical point of view. I understand the difference in the formula.

8. I didn't understand X11. (Classical method in Canada)
9. Different components in the last column: random, irregular and remainder.
10. The example in X_11 (line 205) has the same function as X_13ARIMA_SEATS (line 238)
11. I cant tell which one we should use for each situation
















