---
title: "Chapter 04 - fpp3"
author: "Leticia"
date: '2022-10-28'
output: html_document
---

# Libraries
```{r Libraries}
library(fpp3)
library(GGally)

```

# Simple Stats
```{r}
tourism %>%
  features(Trips, list(mean = mean)) %>%
  arrange(mean)
```

```{r}
tourism %>% features(Trips, quantile)
```

```{r}
tourism %>% features(Trips, feat_acf)

```

# STL Features

FS (Seasonal Strength)  close to 0 exhibits almost no seasonality, while a series with strong seasonality will have FS close to 1 because Var(Rt) will be much smaller than Var(St+Rt).

```{r}
tourism %>%
  features(Trips, feat_stl)
```
# Plotting Seasonal Strength

Seasonal strength vs trend strength for all tourism series.
```{r}
tourism %>%
  features(Trips, feat_stl) %>%
  ggplot(aes(x = trend_strength, y = seasonal_strength_year,
             col = Purpose)) +
  geom_point() +
  facet_wrap(vars(State))
```

# The most seasonal series in the Australian tourism data
```{r}
tourism %>%
  features(Trips, feat_stl) %>%
  filter(
    seasonal_strength_year == max(seasonal_strength_year) #Filter here
  ) %>%
  left_join(tourism, by = c("State", "Region", "Purpose")) %>%
  ggplot(aes(x = Quarter, y = Trips)) +
  geom_line() +
  facet_grid(vars(State, Region, Purpose))
```
# All feasts features (48 features)

```{r}
tourism_features <- tourism %>%
  features(Trips, feature_set(pkgs = "feasts"))
tourism_features

```
# Pairwise plots

Pairwise plots of all the seasonal features for the Australian tourism data
```{r}
library(glue)
tourism_features %>%
  select_at(vars(contains("season"), Purpose)) %>%
  mutate(
    seasonal_peak_year = seasonal_peak_year +
      4*(seasonal_peak_year==0),
    seasonal_trough_year = seasonal_trough_year +
      4*(seasonal_trough_year==0),
    seasonal_peak_year = glue("Q{seasonal_peak_year}"),
    seasonal_trough_year = glue("Q{seasonal_trough_year}"),
  ) %>%
  GGally::ggpairs(mapping = aes(colour = Purpose))
```
# Principal components of tourism

A plot of the first two principal components, calculated from the 48 features of the Australian quarterly tourism data.
```{r}
library(broom)
pcs <- tourism_features %>%
  select(-State, -Region, -Purpose) %>%
  prcomp(scale = TRUE) %>%
  augment(tourism_features)
pcs %>%
  ggplot(aes(x = .fittedPC1, y = .fittedPC2, col = Purpose)) +
  geom_point() +
  theme(aspect.ratio = 1)
```



#Outliers

```{r}
outliers <- pcs %>%
  filter(.fittedPC1 > 10) %>%
  select(Region, State, Purpose, .fittedPC1, .fittedPC2)
outliers
```
#Plot Outliers
```{r}
outliers %>%
  left_join(tourism, by = c("State", "Region", "Purpose")) %>%
  mutate(
    Series = glue("{State}", "{Region}", "{Purpose}",
                  .sep = "\n\n")
  ) %>%
  ggplot(aes(x = Quarter, y = Trips)) +
  geom_line() +
  facet_grid(Series ~ ., scales = "free") +
  labs(title = "Outlying time series in PC space")
```

# Exercise
## 2 
Use GGally::ggpairs() to look at the relationships between the STL-based features for the holiday series in the tourism data. Change seasonal_peak_year and seasonal_trough_year to factors, as shown in Figure 4.3. Which is the peak quarter for holidays in each state?


```{r}
library(glue)
tourism_features %>%
  select_at(vars(contains("season"), Purpose)) %>%
  mutate(seasonal_peak_year=as.factor(seasonal_peak_year)) %>%
  mutate(seasonal_trough_year=as.factor(seasonal_trough_year)) %>% 
  mutate(
    seasonal_peak_year = seasonal_peak_year +
      4*(seasonal_peak_year==0),
    seasonal_trough_year = seasonal_trough_year +
      4*(seasonal_trough_year==0),
    seasonal_peak_year = glue("Q{seasonal_peak_year}"),
    seasonal_trough_year = glue("Q{seasonal_trough_year}"),
  ) %>%
  GGally::ggpairs(mapping = aes(colour = Purpose))
```













# Questions
- How to identify the years on line 41
- Hoe entropy and timeseries are related exactly? (lack of order or predictability; gradual decline into disorder)

