---
title: "Chapter5"
author: "Leticia"
date: '2022-11-04'
output: html_document
---

# Libraries
```{r}
library(fpp3)
```

# Data preparation (tidy)
```{r}
gdppc <- global_economy %>%
  mutate(GDP_per_capita = GDP / Population)
  
  
  gdppc
```
# Plot the data (visualise)
```{r}
gdppc %>%
  filter(Country == "Sweden") %>%
  autoplot(GDP_per_capita) +
  labs(y = "$US", title = "GDP per capita for Sweden")
```
```{r}

gdppc #%>%
  filter(Country == "Bangladesh") %>%
  autoplot(GDP_per_capita) +
  labs(y = "$US", title = "GDP per capita for Bangladesh")
  
  
```

```{r}
# Re-index based on trading days
gdppc %>%
  filter(Country == "Bangladesh", year(Date) >= 1990) #%>%
  mutate(day = row_number()) %>%
  update_tsibble(index = day, regular = TRUE)
# Filter the year of interest
Bangladesh_1990 <- Bangladesh_2015 %>% filter(year(Date) == 1990)
# Fit the models
Bangladesh_fit <- Bangladesh_2015 %>%
  model(
    Mean = MEAN(Close),
    `Naïve` = NAIVE(Close),
    Drift = NAIVE(Close ~ drift())
  )
# Produce forecasts for the trading days in January 2016
Bangladesh_jan_2016 <- Bangladesh_1990 %>%
  filter(yearmonth(Date) == yearmonth("2016 Jan"))
Bangladesh_fc <- Bangladesh_fit %>%
  forecast(new_data = Bangladesh_jan_2016)
# Plot the forecasts
google_fc %>%
  autoplot(Bangladesh_2015, level = NULL) +
  autolayer(Bangladesh_jan_2016, Close, colour = "black") +
  labs(y = "$US",
       title = "GDP per capita for Bangladesh",
       subtitle = "(Jan 2015 - Jan 2016)") +
  guides(colour = guide_legend(title = "Forecast"))
```



# Define a model (specify)
```{r}
TSLM(GDP_per_capita ~ trend()) #Time series linear model
```

# Train the model (estimate)
```{r}
fit <- gdppc %>%
  model(trend_model = TSLM(GDP_per_capita ~ trend()))
```
# Check model performance (evaluate)
How well it performed? Make some Diagnostics.

# Produce forecasts (forecast)

```{r}
fit %>% forecast(h = "3 years")
```

```{r}
fit %>%
  forecast(h = "3 years") %>%
  filter(Country == "Sweden") %>%
  autoplot(gdppc) +
  labs(y = "$US", title = "GDP per capita for Sweden")
```
# 5.2 Some simple forecasting methods

```{r}
bricks <- aus_production %>%
  filter_index("1970 Q1" ~ "2004 Q4") %>%
  select(Bricks)

bricks
```
Mean method

```{r}
bricks %>% model(MEAN(Bricks))
autoplot(bricks)
```
Naïve method
```{r}
bricks %>% model(NAIVE(Bricks))
autoplot(bricks)
```
# Seasonal naïve method
```{r}
bricks %>% model(SNAIVE(Bricks ~ lag("year")))
autoplot(bricks)
```
# Drift method

```{r}
bricks %>% model(RW(Bricks ~ drift()))
autoplot(bricks)
```

```{r}
# Set training data from 1992 to 2006
train <- aus_production %>%
  filter_index("1992 Q1" ~ "2006 Q4")
# Fit the models
beer_fit <- train %>%
  model(
    Mean = MEAN(Beer),
    `Naïve` = NAIVE(Beer),
    `Seasonal naïve` = SNAIVE(Beer)
  )
# Generate forecasts for 14 quarters
beer_fc <- beer_fit %>% forecast(h = 14)
# Plot forecasts against actual values
beer_fc %>%
  autoplot(train, level = NULL) +
  autolayer(
    filter_index(aus_production, "2007 Q1" ~ .),
    colour = "black"
  ) +
  labs(
    y = "Megalitres",
    title = "Forecasts for quarterly beer production"
  ) +
  guides(colour = guide_legend(title = "Forecast"))
```

# Example: Google’s daily closing stock price
```{r}
# Re-index based on trading days
google_stock <- gafa_stock %>%
  filter(Symbol == "GOOG", year(Date) >= 2015) %>%
  mutate(day = row_number()) %>%
  update_tsibble(index = day, regular = TRUE)
# Filter the year of interest
google_2015 <- google_stock %>% filter(year(Date) == 2015)
# Fit the models
google_fit <- google_2015 %>%
  model(
    Mean = MEAN(Close),
    `Naïve` = NAIVE(Close),
    Drift = NAIVE(Close ~ drift())
  )
# Produce forecasts for the trading days in January 2016
google_jan_2016 <- google_stock %>%
  filter(yearmonth(Date) == yearmonth("2016 Jan"))
google_fc <- google_fit %>%
  forecast(new_data = google_jan_2016)
# Plot the forecasts
google_fc %>%
  autoplot(google_2015, level = NULL) +
  autolayer(google_jan_2016, Close, colour = "black") +
  labs(y = "$US",
       title = "Google daily closing stock prices",
       subtitle = "(Jan 2015 - Jan 2016)") +
  guides(colour = guide_legend(title = "Forecast"))
```
# 5.3 Fitted values and residuals

.fitted contains the fitted values;
.resid contains the residuals;
.innov contains the “innovation residuals” which, in this case, are identical to the regular residuals

```{r}
augment(beer_fit) # Augment is for visualizing residuals
```
5.4 Residual diagnostics
1. The innovation residuals are uncorrelated. If there are correlations between innovation residuals, then there is information left in the residuals which should be used in computing forecasts.
2. The innovation residuals have zero mean. If they have a mean other than zero, then the forecasts are biased.
3.  innovation residuals have constant variance. This is known as “homoscedasticity”.
4. The innovation residuals are normally distributed.

Example: Forecasting Google daily closing stock prices

```{r}
autoplot(google_2015, Close) +
  labs(y = "$US",
       title = "Google daily closing stock prices in 2015")
```

```{r}
aug <- google_2015 %>%
  model(NAIVE(Close)) %>%
  augment()
autoplot(aug, .innov) +
  labs(y = "$US",
       title = "Residuals from the naïve method")
```

```{r}
aug <- google_2015 %>%
  model(NAIVE(Close)) %>%
  augment()
autoplot(aug, .innov) +
  labs(y = "$US",
       title = "Residuals from the naïve method")
```

```{r}
aug %>%
  ggplot(aes(x = .innov)) +
  geom_histogram() +
  labs(title = "Histogram of residuals")
```


```{r}
aug %>%
  ACF(.innov) %>%
  autoplot() +
  labs(title = "Residuals from the naïve method")
```

```{r}
google_2015 %>%
  model(NAIVE(Close)) %>%
  gg_tsresiduals()
```


# Portmanteau tests for autocorrelation
```{r}
aug %>% features(.innov, box_pierce, lag = 10, dof = 0)
```


```{r}
aug %>% features(.innov, ljung_box, lag = 10, dof = 0)

```

```{r}
fit <- google_2015 %>% model(RW(Close ~ drift()))
tidy(fit)
```

```{r}
augment(fit) %>% features(.innov, ljung_box, lag=10, dof=1)

```


# How to intentify bias and correlation in time series? What is the difference, because to me it seems like is the same thing.
# How can we plot Augment?








